---
title: '4.10 The Cox model fitting'
author: "Hening Cui"
date: "4/4/2023"
output: html_document
---
# Preparing Data
## loading library
```{r, warning=FALSE}
library(Seurat)
library(spacexr)
library(tidyverse)
library(spatstat)
library(nlme)
library(sp)
```

## loading dataset
```{r}
load(file= "/Users/cuihening/Desktop/shikun/seurat/robjs/chicken_visium.4.Robj")
```

```{r}
TM <- FetchData(object=chicken_visium,vars=c('TMSB4X','ident', 'orig.ident')) 
tmsb4x=cbind(cell = rownames(TM), TM)
```


# Try first with D4 data

## Prepare data

D4 coordinate

```{r}
coords_d4 <- GetTissueCoordinates(chicken_visium, image= 'slice1')
colnames(coords_d4) <- c("x", "y")
coords_d4 <- cbind(cell = rownames(coords_d4),coords_d4)
```



D4 dataset and 
```{r}
D4 = tmsb4x %>% 
  filter(orig.ident=='D4') %>% 
  merge(coords_d4, by="cell") 
```


```{r}
X <- D4$x
Y <- D4$y
xrange <- range(X, na.rm=T)
yrange <- range(Y, na.rm=T)
D4_samppp =  ppp(X,Y,c(380, 484),c(268, 386), marks=D4["TMSB4X"])
```


```{r}
plot(unmark(D4_samppp))
```

```{r}
plot(density(D4_samppp))
```





```{r}
hist(marks(D4_samppp), main="D4-Distribution of TMSB4X exprssion")
```

```{r}
D4_sampty =  ppp(X,Y,c(380, 484),c(268, 386),marks = D4$ident)
D4cell_type= split(D4_sampty)
D4_ty2 = D4cell_type$'2'
```

```{r}
D4_ty2df = D4 %>% filter(ident==2)
LMM <- lme (TMSB4X ~ x+y, random = ~1 | cell,  data = D4_ty2df, method='REML')
summary(LMM)
```


```{r}
cf=coef(LMM)
TMfunction <- function(x,y){ cf$`(Intercept)`	-0.00007802*x + 0.00011638*y}
D4_fit =  kppm(D4_ty2 ~TMSB4X, "LGCP", data = D4_ty2, covariates=list(TMSB4X=TMfunction))
```

```{r}
summary(D4_fit)
plot(D4_fit)
```


## EDA
Build D4 data ppp project 
```{r unmark data}
X <- D4$x
Y <- D4$y
xrange <- range(X, na.rm=T)
yrange <- range(Y, na.rm=T)
D4_tm =  ppp(X,Y,xrange,yrange,marks = D4$TMSB4X)
par(mfrow=c(2,2))
plot(unmark(D4_tm),main="D4 only point pattern")
plot(D4_tm,main="TMSB4X marker spatial plot")
hist(marks(D4_tm), main="D4-Distribution of TMSB4X exprssion")
D4_ty =  ppp(X,Y,xrange,yrange,marks = D4$ident)
plot(D4_ty, main="D4 cell type distribution")
```




Different cell type density plot

```{r}
D4cell_type= split(D4_ty)
plot(D4cell_type[0:9], main = "D4 cell location by type")
plot(density(D4cell_type[0:9]), main = "D4 Densities of cell type")
```

## For type 2 cell fitting model

Type 2 cell data

```{r}
D4_ty2 = D4cell_type$'2'
```



Fitting LMM model for the TMSB4
```{r}
D4_ty2df = D4 %>% filter(ident==2)
LMM <- lme (TMSB4X ~ x+y, random = ~1 | cell,  data = D4_ty2df, method='REML')
summary(LMM)
```

Fitting Cox spatial model

```{r}
cf=coef(LMM)
TMfunction <- function(x,y){ cf$`(Intercept)`	-0.00007802*x + 0.00011638*y}
D4_fit =  kppm(unmark(D4_ty2) ~TMSB4X, "LGCP", data = D4_ty2, covariates=list(TMSB4X=TMfunction))
```


```{r}
summary(D4_fit)
plot(D4_fit)
```


# Pseudotime analysis

Take type2 as an example.

## Type 2 all time
```{r}
Type2 = tmsb4x %>% 
  filter(ident==2) %>% 
  dplyr::rename(time = orig.ident)
```


The function to get coordinate
```{r}
getcoor=function(slice, time){
coords <- GetTissueCoordinates(chicken_visium, image= slice)
colnames(coords) <- c("x", "y")
coords <- cbind(cell = rownames(coords),coords)
coords$time = time
return(coords)
}
```

All the time point coordinates
```{r}
coords_d4 = getcoor('slice1','D4')
coords_d10 = getcoor('slice1_D10-C1','D10')
coords_d7 = getcoor('slice1_D7-B1','D7')
coords_d14 = getcoor('slice1_D14-D1','D14')
```

```{r}
all_coor= tmsb4x %>% 
  dplyr::rename(time = orig.ident) %>% 
  merge(coords_d4,by="cell") %>% 
  rbind(merge(Type2, coords_d7,by="cell")) %>% 
  rbind(merge(Type2, coords_d10,by="cell")) %>% 
  rbind(merge(Type2, coords_d14,by="cell")) %>% 
  mutate(time = factor(time.x))
```
Integrate to the type 2 dataset


```{r}
ty2_coor= Type2 %>% 
  merge(coords_d4,by="cell") %>% 
  rbind(merge(Type2, coords_d7,by="cell")) %>% 
  rbind(merge(Type2, coords_d10,by="cell")) %>% 
  rbind(merge(Type2, coords_d14,by="cell")) %>% 
  mutate(time = factor(time.x))
```

```{r}
D7_ty2 = ty2_coor %>% 
  filter(time=='D7') 
```


```{r}
X <- D7_ty2$x
Y <- D7_ty2$y
xrange <- range(X, na.rm=T)
yrange <- range(Y, na.rm=T)
D7_samppp =  ppp(X,Y,c(255, 480),c(267,470), marks=D7_ty2["TMSB4X"])
```


```{r}
plot(unmark(D7_samppp))
```
```{r}
LMM <- lme (TMSB4X ~ x+y, random = ~1 | cell,  data = D7_ty2, method='REML')
summary(LMM)
```


```{r}
cf=coef(LMM)
TMfunction <- function(x,y){ cf$`(Intercept)`	+ 0.00000425*x - 	-0.00000400*y}
D7_fit =  kppm(unmark(D7_samppp) ~TMSB4X, "LGCP", data = D7_ty2, covariates=list(TMSB4X=TMfunction))
```

```{r}
summary(D7_fit)
plot(D7_fit)
```


```{r}
D10_ty2 = ty2_coor %>% 
  filter(time=='D10') 
```


```{r}
X <- D10_ty2$x
Y <- D10_ty2$y
xrange <- range(X, na.rm=T)
yrange <- range(Y, na.rm=T)
D10_samppp =  ppp(X,Y,c(193,547),c(320,538), marks=D10_ty2["TMSB4X"])
```


```{r}
plot(unmark(D10_samppp))
```

```{r}
LMM <- lme (TMSB4X ~ x+y, random = ~1 | cell,  data = D10_ty2, method='REML')
summary(LMM)
```


```{r}
cf=coef(LMM)
TMfunction <- function(x,y){ cf$`(Intercept)`	+ 0.00000311*x  	-0.00000870*y}
D10_fit =  kppm(unmark(D10_samppp) ~TMSB4X, "LGCP", data = D10_ty2, covariates=list(TMSB4X=TMfunction))
```

```{r}
summary(D10_fit)
plot(D10_fit)
```

```{r}
D14_ty2 = ty2_coor %>% 
  filter(time=='D14') 
```


```{r}
X <- D14_ty2$x
Y <- D14_ty2$y
xrange <- range(X, na.rm=T)
yrange <- range(Y, na.rm=T)
D14_samppp =  ppp(X,Y,xrange, yrange, marks=D14_ty2["TMSB4X"])
```


```{r}
plot(unmark(D14_samppp))
```

Build the ppp object

```{r}
buildppp=function(dataset,marker){
X <- dataset$x
Y <- dataset$y
xrange <- range(X, na.rm=T)
yrange <- range(Y, na.rm=T)
ppppro =  ppp(X,Y,xrange,yrange,marks = dataset[marker])
return(ppppro)
}
```

```{r}
allppp=buildppp(all_coor, 'time')
all_time =split(allppp)
plot(all_time[0:4], main = "cell location")
```

## EDA
```{r}
type2ppp =  buildppp(ty2_coor, 'TMSB4X')
hist(marks(type2ppp), main="Type 2Distribution of TMSB4X exprssion")
```



```{r, fig.height=12, fig.width=20}
type2_alltime =  buildppp(ty2_coor, 'time')
ty2_time= split(type2_alltime)
plot(ty2_time[0:4], main = "Type 2 cell location by type")
plot(density(ty2_time[0:4]), main = "Type 2 cell Densities across time")
```


## Model fit

```{r}
fitcox = function(data, tim, ppplist){
  dataset=  data %>% filter(time == tim)
  LMM = lme (TMSB4X ~ x+y, random = ~1 | cell,  data = dataset, method='REML')
  ppppro = ppplist[[tim]]
  cf=coef(LMM)
  TMfunction <- function(x,y){ cf$`(Intercept)`	-cf$x*x + cf$y*y}
  cox_fit =  kppm(unmark(ppppro) ~TMSB4X, "LGCP", data = ppppro, covariates=list(TMSB4X=TMfunction))
  return(cox_fit)}
```


```{r}
D4 = fitcox(ty2_coor,'D4',ty2_time)
summary(D4)
D7 = fitcox(ty2_coor,'D7',ty2_time)
summary(D7)
D10 = fitcox(ty2_coor,'D10',ty2_time)
summary(D10)
D14 = fitcox(ty2_coor,'D14',ty2_time)
summary(D14)
```

```{r}
plot(D4, main="D4 type 2")
plot(D7, main="D7 type 2")
plot(D10, main="D10 type 2")
```


```{r}
plot(D14)
```

```{r}

```






