---
title: "Temporal Cox model"
author: "Hening Cui"
date: "5/9/2023"
output: 
  html_document:
   code_folding: hide
---
```{r, warning=FALSE, message=FALSE}
library(Seurat)
library(spacexr)
library(tidyverse)
library(spatstat)
library(nlme)
library(lgcp)
library(CAinterprTools)
library(sf)
library(concaveman)
library(ggplot2)
```

# loading dataset
```{r}
load(file= "/Users/cuihening/Desktop/shikun/seurat/robjs/chicken_visium.4.Robj")
```

```{r}
TM <- FetchData(object=chicken_visium,vars=c('TMSB4X','ident', 'orig.ident')) 
tmsb4x=cbind(cell = rownames(TM), TM) %>% 
  dplyr::rename(time = orig.ident) 
tmsb4x$is2=as.factor(ifelse(tmsb4x$ident==2,1,0))
```

The function to get coordinate
```{r}
getcoor=function(slice, time){
coords <- GetTissueCoordinates(chicken_visium, image= slice)
colnames(coords) <- c("x", "y")
coords <- cbind(cell = rownames(coords),coords)
coords$time = time
return(coords)
}
```

```{r}
coords_d4 = getcoor('slice1','D4')
coords_d10 = getcoor('slice1_D10-C1','D10')
coords_d7 = getcoor('slice1_D7-B1','D7')
coords_d14 = getcoor('slice1_D14-D1','D14')
```

Function for constuct ppp

```{r}
buildppp=function(dataset,markers){
X <- dataset$x
Y <- dataset$y
xrange <- range(X, na.rm=T)
print(xrange)
yrange <- range(Y, na.rm=T)
print(yrange)
ppppro =  ppp(X,Y,xrange,yrange,marks=markers)
return(ppppro)
}
```

# EDA 

## D4

All day4 sample

```{r}
D4 = tmsb4x %>% 
  filter(time=='D4') %>% 
  merge(coords_d4, by="cell")

D4_samppp = ppp(D4$x,D4$y, c(0, 450),c(0,550),marks=D4$is2)

plot(D4_samppp)
```

only sample with type2


```{r}
#create dataset
D4 = tmsb4x %>% 
  filter(time=='D4') %>% 
  merge(coords_d4, by="cell")  %>% 
  filter(x >= 375 &  y>=265 )  %>% 
  filter(!(x >= 375 & x < 385 & y < 300)) %>% 
  dplyr::rename(Y=x, X=y) %>% 
  dplyr::rename(x=X, y=Y) %>% 
  mutate(x=x-138, y=y-208) %>% 
  mutate(time=4)
summary(D4)

#generate window
xym= D4 %>% select(c("x","y")) 
pts <- st_as_sf(xym, coords=c('x','y'))
conc <- concaveman(pts)
ggplot() +
  geom_sf(data = conc, fill = NA)

d4_owin= as.owin(as_Spatial(conc))
plot(d4_owin)

#creat ppp project
D4_samppp =  ppp(D4$x,D4$y, window = d4_owin, marks=D4$is2)
```

For type 2 cell
```{r}
D4_ty2=D4 %>% 
  filter(is2==1)

D4_ty2_owin=ppp(D4_ty2$x,D4_ty2$y, window = d4_owin)
```

EDA
```{r}
par(mfrow=c(1,2))
plot(D4_samppp, pch = 20,show.window=FALSE,cols=c("blue", "green"), legend=FALSE, main="D4 point",markscale = 0.1)
plot(density(D4_ty2_owin), main="D4 density")
```

```{r}
pixel=pixellate.ppp(D4_ty2_owin,eps=c(5,5),fractional=TRUE)
plot(pixel)
```






### D7


All day7 sample

```{r}
D7 = tmsb4x %>% 
  filter(time=='D7') %>% 
  merge(coords_d7, by="cell")

summary(D7)

D7_samppp = ppp(D7$x,D7$y, c(0, 450),c(0,550),marks=D7$is2)

plot(D7_samppp)
```
extract one sample

```{r}
D7 = tmsb4x %>% 
  filter(time=='D7') %>% 
  merge(coords_d7, by="cell")  %>% 
  filter(x <= 255 &  y>=285 )  %>% 
  rename(Y=x, X=y) %>% 
  rename(x=X, y=Y) %>% 
  mutate(x=x-150, y=300-y) %>% 
  mutate(time=7)
summary(D7)
```

```{r}
xym= D7 %>% select(c("x","y")) 
pts <- st_as_sf(xym, coords=c('x','y'))
conc <- concaveman(pts)
ggplot() +
  geom_sf(data = conc, fill = NA)
```



```{r}
d7_owin= as.owin(as_Spatial(conc))
plot(d7_owin)
D7_samppp =  ppp(D7$x,D7$y, d7_owin , marks=D7$is2)
plot(D7_samppp)
```
```{r}
D7_samppp =  ppp(D7$x,D7$y, window = d7_owin, marks=D7$is2)
```

```{r}
D7_ty2=D7 %>% 
  filter(is2==1)
D7_ty2p=ppp(D7_ty2$x,D7_ty2$y, window = d7_owin)
```

```{r, fig.height=10, fig.width=12}
par(mfrow=c(1,2))
plot(D7_samppp, pch = 20,show.window=FALSE, cols=c("blue", "green"), legend=FALSE, main="D7 point")
plot(density(D7_ty2p), main="D7 density")
```

```{r}
pixel=pixellate.ppp(D7_ty2p,eps=c(4,4),fractional=TRUE)
plot(pixel)
```



### D10

All D10 sample
```{r}
D10 = tmsb4x %>% 
  filter(time=='D10') %>% 
  merge(coords_d10, by="cell")

D10_samppp =ppp(D10$x,D10$y, c(0, 550),c(0,550), marks=D10$is2)
plot(D10_samppp)
```

Try to extract one sample

```{r}
D10 = tmsb4x %>% 
  filter(time=='D10') %>% 
  merge(coords_d10, by="cell")  %>% 
  filter(y < 340)  %>% 
  rename(Y=x, X=y) %>% 
  rename(x=X, y=Y) %>% 
  mutate(x=x-67, y=y-128) %>% 
  mutate(time=10)
summary(D10)
D10_samppp =ppp(D10$x,D10$y, c(0, 450),c(0,550), marks=D10$is2)
plot(D10_samppp)
```

```{r}
xym= D10 %>% select(c("x","y")) 
pts <- st_as_sf(xym, coords=c('x','y'))
conc <- concaveman(pts)
ggplot() +
  geom_sf(data = conc, fill = NA)
```

```{r}
d10_owin= as.owin(as_Spatial(conc))
plot(d10_owin)
```
```{r}
D10_samppp =ppp(D10$x,D10$y, window = d10_owin, marks=D10$is2)
```

```{r}
D10_ty2=D10 %>% 
  filter(is2==1)
D10_ty2p=ppp(D10_ty2$x,D10_ty2$y, window = d10_owin)
```

```{r, fig.height=10, fig.width=12}
par(mfrow=c(1,2))
plot(D10_samppp, pch = 20,show.window=FALSE, cols=c("blue", "green"), legend=FALSE, main="D10 point")
plot(density(D10_ty2p), main="D10 density")
```
```{r}
D10_fit =  kppm(D10_ty2p ~1, "LGCP")
D10_fit
```

## D14
```{r}
D14 = tmsb4x %>% 
  filter(time=='D14') %>% 
  merge(coords_d14, by="cell")  %>% 
  rename(Y=x, X=y) %>% 
  rename(x=X, y=Y) %>% 
  mutate(x=x-144, y=y-78) %>% 
  mutate(time=14)
summary(D14)
```

```{r}
xym= D14 %>% select(c("x","y")) 
pts <- st_as_sf(xym, coords=c('x','y'))
conc <- concaveman(pts)
ggplot() +
  geom_sf(data = conc, fill = NA)
```

```{r}
d14_owin= as.owin(as_Spatial(conc))
plot(d14_owin)
```

```{r}
D14_samppp =  ppp(D14$x,D14$y, window = d14_owin, marks=D14$is2)
```



```{r}
D14_ty2=D14 %>% 
  filter(is2==1)
D14_ty2p=ppp(D14_ty2$x,D14_ty2$y, window = d14_owin)
```

```{r, fig.height=10, fig.width=12}
par(mfrow=c(1,2))
plot(D14_samppp, pch = 20,show.window=FALSE, cols=c("blue", "green"), legend=FALSE, main="D14 point")
plot(density(D14_ty2p), main="D14 density")
```

```{r}
D14_fit =  kppm(D14_ty2p ~1, "LGCP")
D14_fit
```

## LGCP model
try fit LGCP model

```{r}
D4_fit =  kppm(D4_ty2_owin ~x+y, "LGCP")
D4_fit
```


```{r}
plot(predict.dppm(D4_fit))
```

## Spatstat replicate model
Create the list of all point
```{r}
time_ty2=solist(D4_ty2_owin,D7_ty2p, D10_ty2p, D14_ty2p)
time_ty2
sapply(time_ty2, npoints)
```

```{r}
plot.solist(time_ty2, main="The ty2 cell", main.panel=letters[1:4], legend=FALSE, equal.scales = TRUE, halign=TRUE, valign=TRUE)
```

Turn list into hyperframe
```{r}
hyoertime <- hyperframe(cell=time_ty2)
hyoertime$time=c(4,7,10,14)
hyoertime$sub=c(1,1,1,1)
```

```{r}
plot(hyoertime, quote(plot(density(cell), main=time)), halign=TRUE, valign=TRUE, equal.scale=TRUE, equal.ribbon=TRUE, main = "")
title(main="The density of Type 2 cell")
```

```{r}
try_po=mppm(cell~time+x+ y + x*y + x*time+ y*time + I(x^2)+I(y^2), hyoertime)
try_po
plot(try_po, main="The ty2 cell polynomial",show.window=FALSE, cif=TRUE,halign=TRUE, valign=TRUE,how="image", equal.ribbon=TRUE, cex=3)
```




test the model
```{r}
cdf.test.mppm(try_po, "x")
```



```{r}
sub_time=subfits(try_po)
```
```{r}
diagnose.ppm(sub_time$`1`)
```
Point patterns $y(i), i = 1, \ldots, n$ are independent, conditional on the covariates and the random effects.

$$\lambda_i(u; \beta, \gamma) = \exp\left[B_i(u) + \beta^T x_i(u) + \gamma^T z_i(u)\right]$$
where:
- $\beta$ is a vector of fixed coefficients, i.e., $\beta = (\beta_1, \beta_2, \ldots, \beta_p)^T$
- $\gamma$ is a vector of random coefficients, i.e., $\gamma = (\gamma_1, \gamma_2, \ldots, \gamma_p)^T$
- $B_i(u)$ is a known baseline function ---optional (not inlcude)
- $x_i(u) = (x_{i1}(u), \ldots, x_{ip}(u))^T$ is a vector of covariates --time spatial
- $z_i(u) = (z_i(u), \ldots, z_i(u))^T$ is a vector of random effects --- x vs z  sample z distribution of z
- $1_p$ denotes a vector of ones with length $p$

$$\log L(\beta, \gamma) = \sum_{i=1}^{n} \sum_{y_j \in y(i)} \left[ B(i)(y_j) + \beta^T x(i)(y_j) + \gamma^T z(i)(y_j) \right] - \int_W \exp\left[ B(i)(u) + \beta^T x(i)(u) + \gamma^T z(i)(u) \right] \, du$$
mixed model condition gene|x spatial fx|s f(s) ----fs|x(code calculate)

```{r}
res <- residuals(try_po, type="Pearson")
smor <- with(hyperframe(res=res), Smooth(res, sigma=4))
```

contains the point process residuals for the corresponding point pattern dataset.
```{r}
plot(smor, equal.ribbon=TRUE)
```

```{r}
res <- residuals(try, type="raw")
df <- as.data.frame(hyoertime, warn=FALSE)
df$resid <- sapply(res, integral.msr)
plot(resid ~ time, df)
```


```{r}
qqplot.ppm(sub_time$`1`)
```

```{r}
diagnose.ppm(sub_time$`4`)
qqplot.ppm(sub_time$`4`)
```

```{r}
anova.mppm(try_po, test="Chi")
```


## Extract the predicted value
```{r}
# Create an empty list to store predicted values
predicted_values <- list()

# Loop through the point patterns in the hyperframe
for (i in 1:nrow(hyoertime)) {
  # Extract the point pattern
  pp <- hyoertime$cell[[i]]
  
  # Predict the intensity of the point pattern using the fitted model
  intensity <- predict(try_po, newdata=hyperframe(cell=pp))
  
  # Store the predicted intensity in the list
  predicted_values[[i]] <- intensity
}

predicted_dfs <- lapply(predicted_values, as.data.frame)

```

```{r}
predicd4=as.data.frame.im(fitted_intensities$cif[[1]])
predicd7=as.data.frame.im(fitted_intensities$cif[[2]])
predicd10=as.data.frame.im(fitted_intensities$cif[[3]])
predicd14=as.data.frame.im(fitted_intensities$cif[[4]])
```

```{r}
tmsbpre=predicd4 %>% 
  left_join(D4, by)
```

```{r}
lmm_model <- readRDS("lmm_model.rds")
```


```{r}
fitted_intensities <- predict.mppm(try_po, locations=D4_ty2_owin)
fitted_intensities
```







```{r}
K_functions4 <-  Kinhom(D4_ty2_owin, ratio= TRUE)
K_functions7 <-  Kinhom(D7_ty2p, ratio= TRUE)
K_functions10 <-  Kinhom(D10_ty2p, ratio= TRUE)
K_functions14 <-  Kinhom(D14_ty2p, ratio= TRUE)
```

```{r}
Kcomb=pool(K_functions4,K_functions7,K_functions10,K_functions14)
```

```{r}
try=lgcp.estK(Kcomb)
```

```{r}
try
```



